---
title: "Clade v Classes"
author: "James Hunter"
date: "September 2, 2015"
output: pdf_document
---

This uses the "int resist miami.xlsx" dataset. This data is summarized without formulas and without specific references to how it was summarized onto this dataset. The spreadsheet was prepared in 2009.

Read the data into R
```{r}
suppressPackageStartupMessages(library(RColorBrewer))
cladeclass <- t(read.csv("clade_class_excel.csv", header = FALSE)) 
# transform to put clades as rows
# Remove NT, Total rows
cladeclass <- cladeclass[-(4:5),] 
cladeclass <- 100 * cladeclass # Turn proportions into percentages
colnames(cladeclass) <- c("Zero Class", "One Class", "Two Class", "Three Class")
barplot(cladeclass, beside = TRUE,
                col = brewer.pal(3, "Set1"),
                ylim = c(0, 60),
                xlab = "Number of Classes with Resistance",
                ylab = "% of Cases within Clade Showing Resistance",
                main = "Clades Showing Classes of Resistance",
                legend = c("Clade B", "Clade C", "Clade F"))
rownames(cladeclass) <- c("Clade B", "Clade C", "Clade F")
cladeclass
```

This is indeed the original graph.

# Clades by retroviral classes comparison
With Ester Sabino's spreadsheet, test differences among 3 class and 2 class by clade
```{r cladestest}
clades = factor(c("Clade B", "Clade C", "Clade F"))
# 3 class resistance-- read in 3 class file
cl3bcf <- read.csv("cl3byclade.csv", header = FALSE, stringsAsFactors = FALSE)
years <- c(as.character(2003:2009), "total")
clades <- c("B", "C", "F")
colnames(cl3bcf) <- years
rownames(cl3bcf) <- clades
cl3bcf
cladeb3 <- cl3bcf[1,1:7]
cladec3 <- cl3bcf[2,1:7]
t.test(cladeb3, cladec3)
# 2 class resistance
cl2bcf <- read.csv("cl2byclade.csv", header = FALSE, stringsAsFactors = FALSE)

colnames(cl2bcf) <- years
rownames(cl2bcf) <- clades
cl2bcf
#Difference between B and F
cladeb2 <- cl2bcf[1,1:7]
cladef2 <- cl2bcf[3,1:7]
t.test(cladeb2, cladef2)
# 1 class resistance
cl1bcf <- read.csv("cl1byclade.csv", header = FALSE, stringsAsFactors = FALSE)
colnames(cl1bcf) <- years
rownames(cl1bcf) <- clades
cl1bcf
cladec1 <- cl1bcf[2,1:7]
cladef1 <- cl1bcf[3,1:7]
t.test(cladec1, cladef1)
# 0 class resistance
cl0bcf <- read.csv("cl0byclade.csv", header = FALSE, stringsAsFactors = FALSE)
colnames(cl0bcf) <- years
rownames(cl0bcf) <- clades
cl0bcf
#Difference between C and F
cladec0 <- cl0bcf[2,1:7]
cladef0 <- cl0bcf[3,1:7]
t.test(cladec0, cladef0)


```

