---
title: "Resist Subtype"
author: "James Hunter"
date: "August 27, 2015"
output: pdf_document
---
This is a copy of resistsubtype.R from April 2014
The data come from resistsubtipo.xlsx, one of the original files. This file is obviously derived from Anderson's original data, with the resistances being converted into Sim/Não categories. At the end of this, I will replicate the bar graph from "load Juncao db" file from earlier this week.

Legend:
c2 = 2 class resistance
c3 = 3 class resistance
individual class names followed by "yn" indicate that the variables have been classified as boolean - resistance exists or not in a patient.

```{r }
## Set up R data frame based on resistsubtipo.csv
resistst <- read.csv("resistsubtipo.csv", header = TRUE, 
                     stringsAsFactors = FALSE,
                     na.strings = c("*", "?"))
## Set up resistance types as 1/0
resistst$trnyn <- ifelse(resistst$trn > 0, 1, 0)
resistst$trnnyn <- ifelse(resistst$trnn > 0, 1, 0)
resistst$ipyn <- ifelse(resistst$ip > 0, 1, 0)


## Measure 3 class resistance
# using numeric variables
resistst$c3 <- ifelse(resistst$trn > 0 & resistst$trnn > 0
                      & resistst$ip > 0, 1, 0)
## measure 2 class resistance
# using numeric variables
resistst$c2 <- ifelse(with(resistst,
                           ((trn > 0) + (trnn > 0) + (ip >0)) == 2),
                           1, 0)
## Create names for state of resistances
resNames <- c("Absent", "Present")

## Work only on complete cases (1009 with missing data, 
## of which 994 have stBlast = "BF")

resistcomp <- resistst[complete.cases(resistst),]

#resiststn <- resistst
#resiststn$c3 <- ifelse(resiststn$c3 == 1, "Present", "Absent")
#resiststn$c2 <- ifelse(resiststn$c2 == 1, "Present", "Absent")
## Create tables resistcomp data frame
## Whole data frame
with(resistcomp, table(ano, reg))
with(resistcomp, table(ano, st))
resc3 <- with(resistcomp, table(ano, c3))
resc3
resc2 <- with(resistcomp, table(ano, c2))
resc2
restrn <- with(resistcomp, table(ano, trnyn))
restrn
restrnn <- with(resistcomp, table(ano, trnnyn))
restrnn
resip <- with(resistcomp, table(ano, ipyn))
resip
with(resistcomp, table(st, c3))
with(resistcomp, table(st, c2))

## prepare %'s of c2 for insertion in resist - only years 2001 - 2007
resc2prop <- prop.table(resc2, margin = 1)
resc2prop

require(gmodels)
CrossTable(resistcomp$ano, resistcomp$c2, prop.chisq = FALSE, format = "SPSS")
CrossTable(resistcomp$ano, resistcomp$c3, prop.chisq = FALSE, format = "SPSS")

## axis data with year and n
years <- min(resistcomp$ano) : max(resistcomp$ano)
n <- rep(NA, length(years))
for (i in 1 :length(years)){
   n[i] <- resip[i,1] + resip[i,2]
}
tickvec2 <- rep(NA, length(years))
for (i in 1:length(years)) {
   tickvec2[i] <- paste(years[i], 
                        "\n(",n[i],")", sep = "")
}

## Mount data frame with final proportion data
tables <- c("restrn", "restrnn", "resip", "resc2", "resc3")
resistprop <- as.data.frame(years)
resistprop$trn <-  100 * prop.table(restrn, margin = 1)[,2]
resistprop$trnn <-  100 * prop.table(restrnn, margin = 1)[,2]
resistprop$ip <-  100 * prop.table(resip, margin = 1)[,2]
resistprop$c2 <-  100 * prop.table(resc2, margin = 1)[,2]
resistprop$c3 <-  100 * prop.table(resc3, margin = 1)[,2]

## Graph of resistances
plot(resistprop$trn ~ resistprop$years, type = "l",
     lwd = 2, xaxt = "n", ylim = c(0, 60),
     main = "Percentage of Drug Resistance by Year\n2001 - 2007",
     xlab = "Year(n)",
     ylab = "Percentage ")
lines(resistprop$years, resistprop$trnn, lwd = 2, lty = 2)
lines(resistprop$years, resistprop$ip, lwd = 2, lty = 3)
lines(resistprop$years, resistprop$c2, lwd = 2, lty = 4)
lines(resistprop$years, resistprop$c3, lwd = 2, lty = 5)
legend("top", 
       legend = c("NRTI", "NNRTI", "PI", "2 class", "3 class"),
       lwd = 2,
       lty = seq(1,5), cex = 0.8, 
       ncol = 5)
axis(1, at = resistprop$years, labels = tickvec2)
grid(col = "darkgrey")
```

Bar chart equal to earlier bar chart with data from this dataset - comparison of percentage of resistance with subtype

Measure class resistances in same manner as Anderson data

```{r classresist}
suppressPackageStartupMessages(library(dplyr))
resistcomp <- mutate(resistcomp, numclasses = 
                      (trnyn >0) + (trnnyn > 0) + (ipyn > 0))
resistcomp <- mutate(resistcomp, noclass = numclasses == 0)
resistcomp <- mutate(resistcomp, oneclass = numclasses == 1)
resistcomp <- mutate(resistcomp, twoclass = numclasses == 2)
resistcomp <- mutate(resistcomp, threeclass = numclasses == 3)
table(resistcomp$numclasses) #all clades
# get reduced data set for BCF only
resistbcf <- filter(resistcomp, st %in% c("B", "C", "F"))
resistbcf$st <- factor(resistbcf$st, levels = c("B", "C", "F"))
subtype_table <- table(resistbcf$st)
resistbcftable <- table(resistbcf$st, factor(resistbcf$numclasses))
resistbcfprop <- 100 * prop.table(resistbcftable, 1) # in pct terms
resistbcfprop
```

```{r histos}
library(RColorBrewer)
bcfresist <- barplot(resistbcfprop, beside = TRUE,
                col = brewer.pal(3, "Set1"),
                ylim = c(0, 60),
                xlab = "Number of Classes with Resistance",
                ylab = "% of Cases within Clade Showing Resistance",
                main = "Clades Showing Classes of Resistance",
                legend = c("Clade B", "Clade C", "Clade F"))
grid(col = "black")

```

Now, the original graph:

![original](drug class and resistance.pptx.pdf)

Also, against the graph from earlier in the week (Anderson's spreadsheets):

![Anderson data](Anderson resist clade data.tiff)

Comparison of Anderson's Stanford Clade data with the Brazilian clade assignment.

```{r cladecomp}
load("juncao_data.RData") # load the Anderson data
# from the original plan2 data, get the three subtype classifications
subtypes <- data_frame(stA_Hmmer = plan2$subtipo_Hmmer,
                       stA_Blast = plan2$subtipo_blast,
                       stA_st = plan2$subtipo)
# get reduced set for B,C,F only
subtypes_bcf <- filter(subtypes, stA_st %in% c("B", "C", "F") & 
                          stA_Hmmer %in% c("B", "C", "F") & 
                          stA_Blast %in% c("B", "C", "F"))
subtypes_bcf$stA_Blast <- factor(subtypes_bcf$stA_Blast, levels = c("B", "C", "F"))
subtypes_bcf$stA_st <- factor(subtypes_bcf$stA_st, levels = c("B", "C", "F"))
subtypes_bcf$stA_Hmmer <- factor(subtypes_bcf$stA_Hmmer, levels = c("B", "C", "F"))
```
Table of proportions of clades in Anderson spreadsheets
```{r andersonclade}
stAtable <- table(subtypes_bcf$stA_st)
stAbcfprop <- 100 * stAtable/nrow(subtypes_bcf) # in pct terms
round(stAbcfprop,3)
```

Table of proportions of clades in resistsubtipo spreadsheet

```{r resisttipo}
resistbcf_table <- table(resistbcf$st)
resistbcf_prop <- 100 * resistbcf_table/nrow(resistbcf)
round(resistbcf_prop,3)
t.test(resistbcf_prop, stAbcfprop)

```

T-test shows no difference in distributions of B, C, and F clades coming from the two sets of data. 

Conclusion
----------
Since we have been working consistently with the dataset shown in this report (resistcomp), I would suggest continuing with this and adding this version of the bar chart to the paper. It is consistent with the other results on clades and resistance we have been reporting as suggested by the t-test above. Given everyone's lack of familiarity with the data sets we received last week, I would suggest we focus on these we have been working with more directly.